WSSSPE4-paper-reviewer-response

### 2017-12-18 WSSSPE4_report/WSSSPE4_paper.tex
work in intro - move CFP topics here, reduce description of previous events, move CoC to new section (3), work on response letter to match

The Fourth Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE4)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/}} was
held over 2 1/2 days on 12--14 September 2016 in Manchester, England, with 68 attendees.
This location and date was selected so that WSSSPE4 immediately preceded the
First Research Software Engineers (RSE) Conference.
Previous events in the WSSSPE series include
WSSSPE1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1/}}~\cite{WSSSPE1-pre-report,WSSSPE1},
held in conjunction with the SC13 conference;
WSSSPE1.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}}, a
focused workshop organized jointly with the SciPy
conference\footnote{\url{https://conference.scipy.org/scipy2014/participate/wssspe/}};
WSSSPE2\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2/}}~\cite{WSSSPE2-pre-report,WSSSPE2},
held in conjunction with SC14;
WSSSPE2.1\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe2-1/}}, a
focused workshop organized again jointly with
SciPy\footnote{\url{http://scipy2015.scipy.org/ehome/115969/286469/}};
and WSSSPE3\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe3/}}~\cite{WSSSPE3},
held in Boulder, Colorado, USA.
The first WSSSPE workshop was named
``Working towards
Sustainable Software for Science: Practice and Experiences,'' which remains the meaning
of the WSSSPE group, but the workshops after that were named
``Workshop on Sustainable
Software for Science: Practice and Experiences.'' Together these reflect
that WSSSPE is both a community and a set of workshops.

@@

The WSSSPE4 workshop included multiple mechanisms for participation and
encouraged team building around solutions. WSSSPE4 strongly encouraged participation
of early-career scientists, postdoctoral researchers, graduate students,
early-stage researchers, and those from underrepresented groups,
with funds provided to the conference organizers by the National Science
Foundation (NSF), the Gordon and Betty Moore Foundation, the Alfred P.~Sloan Foundation, and the Software
Sustainability Institute (SSI) to support the travel of potential participants
who would not otherwise be able to attend the workshop. These
funds allowed 29 additional people % to attend and participate. A subset of the
% organizing committee reviewed 
out of 44 applicants which 
% ions for travel support and
% competitively selected the awardees, 
included 11 students and five early-career
researchers.
%  In addition, the travel award subcommittee tried to increase the
% diversity of applicants by forwarding the notice of travel support to
% organizations including PyLadies\footnote{\url{http://www.pyladies.com}},
% Django Girls\footnote{\url{https://djangogirls.org}},
% Women Who Code\footnote{\url{https://www.womenwhocode.com}}, and
% Women in HPC\footnote{\url{http://www.womeninhpc.org}}.
%\todo{did we reach out to any others?}
%
WSSSPE4 also included two professional event organizers/facilitators who helped
plan the workshop agenda,
and who engaged with participants during the workshop.

At the workshop, we introduced a Code of Conduct (CoC).%
\footnote{\label{footnote:CoC}\url{http://wssspe.researchcomputing.org.uk/wssspe4/code-of-conduct/}}
The CoC was conceived for the workshop and a guideline for the community of scientists that WSSSPE
supports, including their personal and online interactions (e.g., on
Twitter, in email lists, in the Slack team). The WSSSPE4 CoC is based on the
FORCE11 conference CoC~\cite{FORCE11:CoC}, which is in turn based on the Code4Lib
CoC~\cite{Code4Lib:CoC}.
The main guidelines of the CoC are:
\begin{quote}
    WSSSPE events are community events intended for networking and collaboration
    as well as learning. We value the participation of every member of the
    community and want all attendees to have an enjoyable and fulfilling
    experience. Accordingly, all attendees are expected to show respect and
    courtesy to other attendees throughout the event and in interactions online
    associated with the event.

@@

The CoC was introduced at the beginning of WSSSPE4, with the CoC subcommittee
and a general email address for reporting concerns or incidents, or
asking questions.  One concern was raised after the first half day, 
% regarding how part of the workshop was being run, 
and we changed the workshop
to address this.

This paper is a summary of a longer report~\cite{WSSSPE4-report} on the events at the workshop and the submitted materials.
The remainder of the paper includes the call for papers (\S\ref{sec:preworkshop}); the WSSSPE mission and vision (\S\ref{sec:mission}); a set of presentations that included a keynote, papers, lighting talks, and a panel (\S\ref{sec:presentations}); and the activities of a set of working groups (\S\ref{sec:WGs}).  One full day of the
workshop was spent with participants in the working groups, which occurred in parallel
with each other.  And each of the working groups left with a plan for how they could move
forward.
This report also mentions the Slack channel created for further discussions (\S\ref{sec:slack}), and it highlights an attendee survey (\S\ref{sec:survey}) before concluding (\S\ref{sec:conclusions}).
%An appendix lists the organizers and program committee members (\S\ref{sec:orgcom}).

WSSSPE4 was based on the work done at previous WSSSPE meetings, but aimed
at producing working groups that better continued working after the workshop ended.
In addition, based on feedback after WSSSPE3, it became clear that attendees
had two distinct motivations:  to make a better future
for research software, and to immediately do better research software development.
This led to the idea of WSSSPE4 being partially divided into two tracks:

Suggested contribution topics included:

\item Training:
taining for developing sustainable software; and 
curriculum for software sustainability

>>>>

The Fourth Workshop on Sustainable Software for Science: Practice and Experiences
(WSSSPE4)\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe4/}} was
held over 2 1/2 days on 12--14 September 2016 in Manchester, England, with 68 attendees.
This location and date was selected so that WSSSPE4 immediately preceded the
First Research Software Engineers (RSE) Conference.
Previous WSSSPE events\footnote{The first WSSSPE workshop was named
``Working towards
Sustainable Software for Science: Practice and Experiences,'' which remains the meaning
of the WSSSPE group, but the workshops after that were named
``Workshop on Sustainable
Software for Science: Practice and Experiences.'' Together these reflect
that WSSSPE is both a community and a set of workshops.}, all of which were in the US, were two general, presentation-focused, one-day workshops held with the SC13 and SC14 conferences~\cite{WSSSPE1-pre-report,WSSSPE1, WSSSPE2-pre-report,WSSSPE2},
two half-day workshops\footnote{\url{http://wssspe.researchcomputing.org.uk/wssspe1-1/}, \url{http://wssspe.researchcomputing.org.uk/wssspe2-1/}} held with SciPy 2014 and 2015 that contained presentations about specific sustainable Python software packages, and a 1-and-1/2-day workshop that included teams that self-assembled and discussed focused software sustainability topics~\cite{WSSSPE3}.
%
Based on the work done at these previous WSSSPE meetings, WSSSPE4 was aimed
at producing working groups that would better continue working after the workshop ended.

Specific topics that were discussed in previous meetings and suggested for WSSSPE4 included:

\item Training:
training for developing sustainable software; and 
curriculum for software sustainability

@@

WSSSPE4 included multiple mechanisms for participation and
encouraged team building around solutions. It strongly encouraged participation
of early-career scientists, postdoctoral researchers, graduate students,
early-stage researchers, and those from underrepresented groups,
with funds provided to the conference organizers by the National Science
Foundation (NSF), the Gordon and Betty Moore Foundation, the Alfred P.~Sloan Foundation, and the Software
Sustainability Institute (SSI) to support the travel of potential participants
who would not otherwise be able to attend the workshop. These
funds allowed 29 additional attendees % to attend and participate. A subset of the
% organizing committee reviewed 
(of 44 applicants), 
% ions for travel support and
% competitively selected the awardees, 
including 11 students and five early-career
researchers.
%  In addition, the travel award subcommittee tried to increase the
% diversity of applicants by forwarding the notice of travel support to
% organizations including PyLadies\footnote{\url{http://www.pyladies.com}},
% Django Girls\footnote{\url{https://djangogirls.org}},
% Women Who Code\footnote{\url{https://www.womenwhocode.com}}, and
% Women in HPC\footnote{\url{http://www.womeninhpc.org}}.
%\todo{did we reach out to any others?}
%
WSSSPE4 also included two professional event organizers/facilitators who helped
plan the workshop agenda,
and who engaged with participants during the workshop.

This paper is a summary of a longer report~\cite{WSSSPE4-report} on the workshop and the submitted materials.
The remainder of the paper includes the call for papers (\S\ref{sec:preworkshop}); the WSSSPE Code of Conduct (\S\ref{sec:CoC}) and mission and vision (\S\ref{sec:mission}); a set of presentations that included a keynote, papers, lighting talks, and a panel (\S\ref{sec:presentations}); and the activities of a set of working groups (\S\ref{sec:WGs}).  One full day of the
workshop was spent with participants in the working groups, which occurred in parallel
with each other.  And each of the working groups left with a plan for how they could move
forward.
This report also mentions the Slack channel created for further discussions (\S\ref{sec:slack}), and it highlights an attendee survey (\S\ref{sec:survey}) before concluding (\S\ref{sec:conclusions}).
%An appendix lists the organizers and program committee members (\S\ref{sec:orgcom}).

Post-WSSSPE3 feedback said that attendees
had two distinct motivations for attending:  to make a better future
for research software, and to immediately do better research software development. Thus,
WSSSPE4 was %partially 
planned as two tracks:

%Suggested contribution topics included:
%
%\begin{itemize}
%\renewcommand{\labelenumi}{\textbf{\theenumi}.}
%\setlength{\rightmargin}{1em}
%
%\item Development and Community:
%best practices for developing sustainable software;
%models for funding specialist expertise in software collaborations;
%software tools that aid sustainability;
%academia/industry interaction;
%refactoring/improving legacy scientific software;
%engineering design for sustainable software;
%metrics for the success of scientific software; and
%adaptation of mainstream software practices for scientific software
%
%\item Professionalization:
%career paths;
%RSE as a brand;
%RSE outside of the UK or Europe; and
%Increase incentives in publishing, funding and promotion for better software
%
%\item Training:
%taining for developing sustainable software; and 
%curriculum for software sustainability
%
%\item Credit:
%making the existing credit and citation ecosystem work better for software;
%future credit and citation ecosystem;
%software contributions as a part of tenure review;
%case studies of receiving credit for software contributions; and
%awards and recognition that encourage sustainable software
%
%\item Software publishing:
%journals and alternative venues for publishing software; and
%review processes for published software
%
%\item Software discoverability/reuse:
%proposals and case studies
%
%\item Reproducibility and testing:
%reproducibility in conferences and journals; and
%best practices for code testing and code review
%
%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Code of Conduct}\label{sec:CoC}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

WSSSPE4 introduced a Code of Conduct (CoC)\footnote{\label{footnote:CoC}\url{http://wssspe.researchcomputing.org.uk/wssspe4/code-of-conduct/}}, which was conceived for the workshop and as guidance for the community of scientists that WSSSPE
supports, including their personal and online interactions (e.g., on
Twitter, in email lists, in Slack). This CoC is based on the
FORCE11 conference CoC~\cite{FORCE11:CoC}, which is in turn based on the Code4Lib
CoC~\cite{Code4Lib:CoC}.
The main guidelines are:
\begin{quote}
    WSSSPE events are community events intended for networking and collaboration
    as well as learning. We value the participation of every member of the
    community and want all attendees to have an enjoyable and fulfilling
    experience. Accordingly, all attendees are expected to show respect and
    courtesy to other attendees throughout the event and in interactions online
    associated with the event.

@@

The CoC was discussed at the beginning of WSSSPE4, with the CoC subcommittee
and a general email address for reporting concerns or incidents, or
asking questions.  One concern was raised after the first half day, 
% regarding how part of the workshop was being run, 
and we changed the workshop
to address this.

### 2017-12-18 WSSSPE4_report/jors-response.tex
work in intro - move CFP topics here, reduce description of previous events, move CoC to new section (3), work on response letter to match

Following these instructions, our initial submission had a body of 13 pages, an appendix of 2 pages, and 3 pages of references (as JORS publishes appendices as separate PDF attachments, and does not typeset them.)

Nonetheless, we agree with the reviewers that this can be reduced, and have done so by removing the appendix and as advised by the reviewers, as further discussed below.  The revised version we are resubmitting has been reduced to \textcolor{red}{11} pages of body, and 3 pages of references, a reduction of \textcolor{red}{5} of the original 18 pages.

@@

>>>>

Following these instructions, our initial submission had 1 title page, a body of 12 pages, an appendix of 2 pages, and 3 pages of references. (Note that JORS publishes appendices as separate PDF attachments, and does not typeset them.)

Nonetheless, we agree with the reviewers that this can be reduced, and have done so by removing the appendix and as advised by the reviewers, as further discussed below.  The revised version we are resubmitting has been reduced to 1 title page, \textcolor{red}{10} pages of body, and 3 pages of references, a reduction of \textcolor{red}{5} of the original 18 pages.

@@

\end{quoting}
We've reduced the list of previous events at the start to make this material more of a summary, and less of a list, and moved a little of Section 2 into Section 1 to make the connection of the events and the motivation for WSSSPE4 more strong, including the CFP topics.  We've also moved the Code of Conduct discussion from the introduction to a new Section (3) to better focus the introduction.
\begin{quoting}
